### 2017-10-05 paper.tex
Project created

\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables

\usepackage[utf8]{inputenc}

\usepackage{hyperref}

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

\usepackage{booktabs}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{tikz}
\usetikzlibrary{positioning, calc}
% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[HRI2018]{ACM/IEEE Human-Robot Conference}{March 2018}{Chicago USA} 
\acmYear{2018}
\copyrightyear{2018}

\acmArticle{4}
\acmPrice{15.00}

\graphicspath{{figs/}}

\usepackage{xspace}
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\etal}{et al.\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\vs}{vs.\xspace}
\newcommand{\resp}{\textit{resp.}\xspace}
\newcommand{\cf}{cf.\xspace}

\title{The Freeplay Sandbox: a Novel Methodology for the Evaluation of Social Robotics}

%\author{\IEEEauthorblockN{Séverin Lemaignan, Emmanuel Senft, Tony Belpaeme}
%\IEEEauthorblockA{
%Centre for Robotics and Neural Systems \\
%Plymouth University \\
%Plymouth, United Kingdom \\
%Email: \tt firstname.lastname@plymouth.ac.uk}
%}

\author{[Hidden for review]}
\orcid{1234-5678-9012}
%\affiliation{%
%  \institution{xxx}
%  \streetaddress{xxx}
%  \city{Dublin} 
%  \state{Ohio} 
%  \postcode{43017-6221}
%}
%\email{trovato@corporation.com}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10010520.10010553.10010554</concept_id>
<concept_desc>Computer systems organization~Robotics</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Robotics}

\keywords{Human-Robot Interaction, Social Robotics, Benchmarking, Evaluation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

Evaluating in a rigorous manner human-robot social interactions is notoriously
difficult~\cite{baxter2016characterising}: studies are either conducted in labs
with constrained protocols to allow for robust measurements and a degree of
replicability, but at the cost of ecological validity. Or on the contrary,
studies are run \emph{in the wild}, which may provide better experimental
realism, but often at the expense of rigorous interaction metrics and replicability.

We introduce in this paper a novel interaction situation designed for ecological
validity while ensuring `good' scientific properties (replicable, clear metrics,
possibility of both autonomous and Wizard-of-Oz robot behaviours).
This task focuses on child-robot interaction in a \emph{sandboxed} free-play
environment. We present hereafter the rationale for this task, its technical and
practical aspects (including the open-source source code), as well as two
large datasets that represent interaction baselines for this task.

\end{abstract}

\maketitle

\section{The difficult evaluation of social interactions}
\label{sec:intro}

Evaluating in a rigorous manner human-robot social interactions is notoriously
difficult~\cite{baxter2016characterising}: studies are either conducted in labs
with constrained protocols to allow for robust measurements and a degree of
replicability, but at the cost of ecological validity. Or on the contrary,
studies are run \emph{in the wild}, which may provide better experimental
realism, but often at the expense of rigorous interaction metrics and replicability.

Existing benchmarks (robocup at home) have issues -- the next challenge for HRI
is to tackle complex, unpredictable situations -> free play

This task is based on free child-robot interactions (free play) in a
\emph{sandboxed} environment: while the interaction is free (children do not
have a particular task to perform, except to play), the activity is scaffolded
by a specific apparatus (a large touchscreen) where the interaction take place.
Capturing the entirety of the interaction is consequently possible and
practical. We present in the remaining of the paper the methodology and
experimental procedure associated to the free-play task, as well as the main
analyses that can be conducted on the collected data (including a new annotation
scheme for social interaction).

\subsection{Studying social interactions}

Studying social interactions requires a social situation that elicits
interactions between the participants.  Such situation is typically scaffolded
by a social task, and consequently, the nature of this task influences in
fundamental ways the kind of interactions that might be observed and analysed.

Because they typically target specific social or cognitive skills in a tightly
controlled manner, traditional socio-cognitive tasks are usually simple,
constrained tasks that do not reflect the complexity and dynamics of real-world
interactions. Conversely, state-of-the-art human-robot interaction aims at
capturing, interpreting and acting upon a rich set of
naturally-occurring social interactions. As such, we argue that further advance
in social human-robot interaction requires us to study and tackle tasks that:

\begin{itemize}
    \item are loosely directed, to maximise natural, non-contrived behaviours;
    \item are long enough and varied enough to elicit a large range of
        interaction situations;
    \item foster rich multi-modal interaction, such as simultaneous speech, gesture, and gaze
        behaviours;
    \item evidence complex social dynamics, such as rhythmic coupling, joint attention,
        implicit turn-taking, a level of unpredictability.
\end{itemize}

The challenge lies in designing a social task that exhibits these features,
while maintaining `good' scientific properties (repeatability, replicability,
robust metrics) as well as good practical properties (not requiring unique or
otherwise very costly experimental environments, not requiring very specific
hardware or robotic platform, easy deployment, short enough experimental
sessions to allow for large groups of participants). So far, our community has
struggled with this endeavour~\cite{baxter2016characterising}.

children can engage in open-ended and
non-directive play situations, yet sufficiently
defined to be reproducible;

\subsection{Background on social play}

The choice of a playful interaction is supported by the wealth of social
situations and social behaviours that \emph{play} elicits. Most of the research
in that field builds on the precursory work of Parten who established five
\emph{stages of play}~\cite{parten1932social}:

\begin{enumerate}
    \item {\bf Solitary (independent) play}: Playing separately from
        others, with no reference to what others are doing.
    \item {\bf Onlooker play}: Watching others play. May engage in
        conversation but not engaged in doing. True focus on the children at
        play.
    \item {\bf Parallel play} (adjacent play, social co-action): Playing
        with similar objects, clearly beside others but not with them (near
        but not with others.)
    \item {\bf Associative play}:  Playing with others without
        organization of play activity. Initiating or responding to
        interaction with peers. 
    \item {\bf Cooperative play}: Coordinating one's behavior with that
        of a peer. Everyone has a role, with the emergence of a sense of
        belonging to a group. Beginning of "team work."
\end{enumerate}

\todo{Should the age be added?}

These five stages of play have been extensively discussed and refined over the
last century, yet remain remarkably widely accepted as such, and to the ears of
researchers in social robotics, they may resonate as an implicit roadmap.

\subsection{Framing: face-to-face dyadic and sandboxed child-robot interactions}

To mitigate the technical challenges presented by free play interactions in
their most unconstrained version (like high dimensionality and uncertainties),
we have developed a \emph{free play sandbox}, Figure~\ref{fig|setup}: a large
touchscreen (60cm $\times$ 33cm) is used as an interactive surface
(\emph{sandtray} paradigm~\cite{baxter2012touchscreen}). Two agents (either a
pair of children or one child and one robot) play together by freely moving on a
2D plane items (characters and blocks, see Figure~\ref{fig|sandbox}). A
background image depicts a generic empty environment, with different symbolic
zones (water, grass, path, bushes...). Additionally, the children can draw on
the background using a tool applying a color on their touches. The players do
not have any particular task to complete, they are simply invited to freely
play.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{The Free-play Sandbox}
\label{sec:freeplay}

\subsection{Task}
description of the task and why it fills a gap in the literature/tools used so
far (provide clearer,  and interesting setup, \emph{easy} to reuse)

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{setup_top}
    %\hspace{1em}
    %\includegraphics[width=0.45\columnwidth]{setup_child_robot_top}
    %\includegraphics[width=0.45\columnwidth]{child-child-env}
    \caption{The free-play social interactions sandbox: two children interact in
    a free-play situation, by drawing and manipulating items on a touchscreen.
    Children are facing each other and sit on cushions. Each child wears a
    bright sports bib, either purple or yellow, to facilitate later
    identification.}

    \label{fig|freeplay}
\end{figure}

We have designed a new experimental task, called the \emph{free-play sandbox},
that is based on free play interactions. Pairs of children (4-8 years old) are
invited to freely draw and interact with items displayed on an interactive
table, without any explicit goal set by the experimenter
(Fig.~\ref{fig|freeplay}).  The task is designed so that children can engage in
open-ended and non-directive play, yet it is sufficiently constrained to be
suitable for recording, and allows the reproduction of social behaviour by an
artificial agent in comparable conditions.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.9\columnwidth]{sandbox}
    \caption{Example of a possible game situation. Items (animals,
    characters...) can be dragged over the whole play area, while the background
    picture can be painted over by picking a colour.}

    \label{fig|sandbox}
\end{figure}

The free-play sandbox follows the sandtray
paradigm~\cite{baxter2012touchscreen}: a large touchscreen (60cm $\times$ 33cm,
with multitouch support) is used as an interactive surface (\emph{sandtray}).
Two children play together by freely moving interactive items
(Fig.~\ref{fig|sandbox}) on the surface. A background image depicts a generic
empty environment, with different symbolic colours (water, grass, beach,
bushes...). By drawing on top of the background picture, the children can change
the environment to their liking. The players do not have any particular task to
complete, they are simply invited to freely play for as long as they want.

\todo{almost identical to 1.3 and mention dataset}

In order to constrain the dataset to a tractable domain, the task is limited to
a dyadic face-to-face interaction.  The resulting interaction space is
nevertheless high-dimensional, with two main foci: the game-related actions
(drawing, manipulating items, telling stories) and the social interactions
(monitoring the other's actions, discussing and agreeing on joint actions,
supportive/antagonistic behaviours, etc.).

\paragraph{Socio-cognitive framing of the task}

focus on abstract socio-cognitive facets (robot
perception and manipulation are simplified); well suited for qualitative and
quantitative analysis using metrics like Słowinski’s Individual Motor Signature
(for behavioural alignment), Anderson's~\cite{anderson2004social} coding of children’s free-play
interactions, with-me-ness (for assessment of co-engagment).

carefully framed: as far as possible, they necessitate only the cognitive
capabilities that they evidence (e.g. if they evidence a purely socio-linguistic
mechanism, they will not mandate complex action capabilities – they may benefit
from it, though),

This point is especially important as it allows an incremental path: a robot
should be able to tackle one or several of these task independently of the
others, and a research team may progressively extend their cognitive models to
incorporate more and more of the cognitive capabilities required by the robot to
address all of the tasks.

\subsection{Applications}
\label{sec:applications}

\paragraph{Child-Child Interaction}

observing and quantifying social interactions.

\paragraph{Child-Robot Interaction}
\label{ssec|CRI}
This Freeplay Sandbox provides the opportunity to explore child-robot
interactions in this open-real world environment as shown in Figure
\ref{fig|freeplay}.

Depending of the focus of the study, two modes of control for the robot are
available. If the interest is on evaluating a specific robot behaviour, the
robot can be autonomously controlled using inputs from the different sensors.
This setup allows to explore the impact of different social behaviours on the
children in an open yet standardised environment.

On the other hand, if the focus is on the child behaviour and the technical
aspect is of a lower importance, the robot can be controlled by a human rather
than an algorithm. This paradigm, where a human teleoperate a robot to interact
with a naive partner is called Wizard of Oz (WoZ) and is used in numerous
studies to explore the psychologic side of HRI \cite{riek2012wizard}.

\paragraph{Deep Learning}

Standardised method to gather data and use deep learning for automated
annotation.

Can be used to explore the impact of social behaviours of the robot: "game
action policy" can be identical, but more social behaviours can be desired

Can be used to explore learning of social action policy for a robot.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Technical Implementation}
\label{sec:impl}

The sandbox is implemented using mainly two frameworks: Qt's QML and
\emph{QtQuick}\footnote{\url{http://doc.qt.io/qt-5/qmlapplications.html}}
for the graphical interface of the game, and the \emph{Robot Operating System}
(ROS) for the modular implementation of the data processing and behaviour
generation pipelines. The graphical interface interacts with the decisional
pipeline over a bidirectional QML-ROS bridge that we have developed for that
purpose\footnote{Source code: \texttt{[hidden for blind review]}}.

Figure~\ref{fig|architecture} presents the software architecture of the sandbox.

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{freeplay-sandbox-architecture}
    \caption{Software architecture of the free-play sandbox. Left (purple) nodes
    are tided to the sandtray (game interface (1) and camera drivers (2)). Nodes in the
    centre (green) implement the behaviour of the robot (play policy (3) and robot
    behaviours (4)). Several helper nodes are available, in particular, segmentation of the
    children drawings into zones (5), A* motion planning for the robot to move
    in-game items (6). Nodes are implemented in
    Python (except for the game interface, developed in QML) and inter-process
    communication relies on ROS. 6D poses are managed and exchanged via ROS TF.}

    \label{fig|architecture}
\end{figure*}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{rviz-sandtray}
    \caption{The free-play sandbox, viewed at runtime within ROS RViz. The
    current background drawing (bottom left) is published on a regular image
    topic. Simple computer vision is used to segment it into zones (visible in
    the central panel). The poses and bounding boxes of the interactive items
    are published as well, and turned into an occupancy map, used to plan the
    robot's arm motion.
    visualisation tool.}
    \label{fig|rviz}
\end{figure}

\subsection{Interactive game}
The interactive game is coded using QML, and possess main background image on
top of which items (animals, humans and objects) can be moved from one part of
the world to another. The children can also use a drawing mode in which some
create strokes of color on a layer between the background and the items. This
increase the openess of the game as symbolic objects can be added by drawing
(eg. Figure~\ref{fig|sandbox}. The game exposes the image of the background
with the drawing in case of change and the positions of each item as a TF frame.

\subsection{Sensors}
Two RGB-D cameras have a fixed position on the Sandtray frame and face one child
each. This fixed position of the cameras permit to extract social features from
the images such as head gaze, gaze, pointing and position in a general
coordinate system usable by each node. In addition of the image, the cameras
posses a microphone to listen and record what the children are saying.

\subsection{Supervisor}
A web interface can be used by an experimenter to manage the meta parameters of
the interaction: demographics, start-stop, task selected, recording...

\subsection{Robot Control}
As stated in section~\ref{ssec|CRI}, a robot can interact in place of one of the
children. This robot can either be autonomous selecting actions based on the
inputs provided by the sensors and the game or be controlled by a human in a
Wizard of Oz fashion.

\paragraph{Autonomous} The current implementation exposes a large number of
information on the game and the state of the child that can be used in the robot
controller. The position of every item is exposed as a TF frame, the background
is segmented in zones of identical colors, social element of the state the
interaction are collected through the RGBD camera and the microphone facing the
child. As visible on Figures~\ref{fig|freeplay} and \ref{fig|rviz}, the camera
covers the head of the child as well as most of the upperbody, and applying
libraries such as DLib and OpenPose, the position of facial feature and skeleton
of the child are extracted and can be used to obtain: head gaze, gaze and
gestures such as pointing. All these inputs can be combined to provide the robot
with more social inputs to test the sociability of a robotic controller and its
impact on the interaction.

The robot's location is obtained by displaying fiducial markers on the
touchscreen before the start of the interaction, so the transformation between
the robot coordinate system and the touchscreen is known. And this robot
location can also be used to identify gazes from the child to the robot.

To make the children believe the robot is moving objects on the touchscreen, we
synchronise a moving pointing gesture of the robot and a series of fake touches
appied on the screen, moving the desired object. Once an object and a goal
position have been selected, a planner generate a path for this image using the
A\mbox{*} algorithm on an occupancy map obtained with the items footprints, then
this plan is sent to a nodes synchronising the actuation on the robot and the
fake touches on the game.

Other actions such as gaze, pointing or speech are also exposed as simple ROS
topics.

\paragraph{Wizard-of-Oz} To allow an experimenter to control the robot, a GUI to control the robot is
provided and presents an identical representation of the state of the game on an
other application which can be used on a tablet for example. The wizard can drag
the objects in a similar fashion as what the child would do on the Sandtray, and
on the release, the robot executes the dragging motion on the Sandtray, moving
an object to a new location. The source code can easily modify to add new
specific buttons to execute other actions, such as having the robot talk to the
child.

\todo{Should I add more in the WoZ GUI in the code side? It's fairly simple in the
current implementation}

\subsection{Records and logs}

Maybe redundant with Data  collection.

\subsection{Tools}

%
%\begin{figure}[h!]
%    \centering
%
%\resizebox{0.9\linewidth}{!}{%
%
%\begin{tikzpicture}[
%    >=latex,
%    node distance=2cm,
%    every edge/.style={draw, very thick},
%    redarrow/.style={fill=Red, draw=Red},
%    greenarrow/.style={fill=GreenYellow, draw=GreenYellow},
%    yellowarrow/.style={fill=BurntOrange, draw=BurntOrange},
%    cmpt/.style={draw, align=center, rounded corners, inner sep=5pt, font=\sf, fill=black!20},
%    label/.style={midway, align=left, font=\scriptsize\sf, fill=white, above,opacity=0,text opacity=1}]
%
%%    \node at (0,0)[cmpt, fill=Cyan!50, text width=2.7cm] (tactile) {tactile interaction \\ manager};
%
%    \node at (0,0) [cmpt, fill=Green!50] (tf) {tf};
%    \node [cmpt, fill=YellowOrange!50, left=of tf] (attention) {attention tracking};
%    \node [above=0.6cm of attention, text width=3cm,align=center] (input) {{\tt
%    /image}, {\tt /camera\_info}, \\ {\tt dlib} {\sf faces model}, \\ {\sf 3D face template}};
%    \node [cmpt, right=of tf] (static) {static pose publishers};
%    \node [above=0.6cm of static, text width=2.7cm,align=center] (staticposes) {\sf locations of \\areas of interest};
%    \node [cmpt, above=0.6cm of tf] (pose) {robot state publisher};
%    \node [cmpt, fill=YellowOrange!50, below=0.6cm of tf] (focus) {focus estimation};
%    \node [below=0.6cm of focus] (output) {\tt /frames\_in\_focus};
%
%    \path (staticposes) edge [->] (static);
%    \path (static) edge [->] node[label,below] {objects poses} (tf);
%    \path (input) edge [->] (attention);
%    \path (attention) edge [->] node[label,below] {child face} (tf);
%    \path (pose) edge [->] node[label,right] {robot pose} (tf);
%    \path (tf) edge [->]  (focus);
%    \path (focus) edge [->]  (output);
%
%\end{tikzpicture}
%}
%    \caption{\small ROS nodes involved in the VFoA estimation (orange nodes were
%    specifically developed for this work).}
%    \label{fig:system}
%\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Canonical procedures for data collection \& analysis}

The section presents \emph{canonical} procedures to acquire data during testing,
to pre-process it, and analyse it. We call them \emph{canonical} because they
are either well integrated into the software pipeline of the sandbox (\eg ROS
integration), represent start-of-the-art techniques, or result from the 
synthesis of existing techniques (this is especially true of the coding scheme
that we introduce for the annotation of social behaviours,
Section~\ref{sec|coding-scheme} below).

However, they do not represent normative procedures. Researchers interested in
reusing the free-play sandbox task for their own research would naturally adapt
and extend these protocols to their own needs. Besides, certain aspects (most
notably, the audio processing) are yet to be properly investigated.

\subsection{Data collection}

Table~\ref{table|datastreams} lists the datastreams that are collected during
the game. By relying on ROS for the data acquisition (and in particular the
\texttt{rosbag} tool), we ensure all the $\approx$10 streams are synchronised,
timestamped, and, where appropriate, come with calibration information (for the
cameras mainly). In our experiments, cameras were configured to stream in qHD
resolution (960$\times$540 pixels) in an attempt to balance high enough
resolution with tractable file size. It results in \emph{bag} files weighting
$\approx$1GB per minute.

In our own experiments, all the data (including up to 5 simultaneous video
streams) was recorded on a single computer (quad core i7-3770T, 8GB RAM)
equipped with a fast 4TB SSD drive. This computer was also running the game
interface on its touch-enabled screen (sandtray), making the whole system
compact and easy to deploy (one single device).

\begin{table}[]
\centering
\caption{List of datastreams typically recorded. Each datastream is timestamped
with a synchronised clock to facilitate later analysis.}
\label{table|datastreams}
\begin{tabular}{@{}ll@{}}
\toprule
\bf Domain  & \bf Type                               \\ \midrule
children    & audio                                  \\
            & face (RGB + depth)                     \\
robot       & full 3D pose                           \\
environment & RGB                                    \\
touchscreen & background drawing (RGB)               \\
            & touches                                \\
            & position and orientation of in-game items         \\
 \multicolumn{2}{l}{static transforms between touchscreen and facial cameras} \\
 \multicolumn{2}{l}{cameras calibration informations}                         \\ \bottomrule
\end{tabular}
\end{table}

\subsection{RGB-D streams analysis}

\paragraph{Face and body pose analysis}

Online processing can be done on the images obtained from the camera. Using new
deep learning libraries such as CMU's OpenPose~\cite{cao2017realtime} and
DLib~(ADDREF), posture and facial landmarks can be obtained online. As the
position of the camera, a potential robot and any object on the game is known,
this landmarks can be mapped to high level behaviours such as poiting or looking
at an object. Additional analysis can be done on the facial landmarks to other
social states, such as main emotion felt by the child.

\subsection{Audio processing}

Similar processing can be applied on the audio stream. Library such as OpenSMILE
provide audio features such as pitch and loudness contour, which inform on the
general state of the child.

As of today, no reliable Speech Recognition Engine exists for
children~\cite{kennedy2017child}, but in the future, the audio should provide
textual information on the requests and comments produced by the child.

\subsection{Game interactions analysis}

Game features are also produced by the different nodes involved in the analaysis
of the game. The Playground segmentation produce a map of the regions based on
the colour which can be used with the positions of the animal to identify from
which zone to which zone an animal has been moved. The relative position of
animal can also indicate if two animals have been moved closer. These relations
and the drawing inform on what high level action the child is doing and can be
used to infer the child's goal or desire.

\subsection{Annotation of Social interactions}
\label{sec|coding-scheme}

Annotating social interaction beyond surface behaviours is generally difficult:
as research scientists, the behaviours we observe are typically the result of a
superposition of complex and non-observable cognitive and emotional states. As
such, these interesting socio-cognitive states can only be indirectly observed,
and their labelling is error prone [TODO: citation needed].

Our aim is to provide insights on the social dynamics, and we have synthesised a
new coding scheme for social interactions that reuse and adapt three established
social scales. Our coding scheme (Figure~\ref{fig|coding-scheme}) looks specifically
at three axis: the level of \emph{task engagement} (that distinguishes between
\emph{focused}, \emph{task oriented} behaviours, and \emph{disengaged} -- yet
sometimes highly social -- behaviours); the level of social engagement (reusing
Parten's stages of play, but at the micro-task level); the social attitude (that
encode attitudes like \emph{supportive}, \emph{aggressive}, \emph{dominant},
\emph{annoyed}, etc.)

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{coding-scheme}
    \caption{The coding scheme used for annotating social interactions occurring
    during free-play episodes. Three main axis are studied: task engagement,
    social engagement and social attitude.}
    \label{fig|coding-scheme}
\end{figure}

\paragraph{Social engagement: Parten's stages of play at micro-level}

\subsection{Video Coding}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Baseline Datasets}
\label{sec:dataset}

We have been using the Freeplay Sandbox task for an initial, large scale, data
collection over a period of 3 months during Spring 2017.

This campaign aimed at (1) extensively evaluating the task itself (would
children engage and exhibit a large range of social dynamics and behaviours?),
(2) making sure the whole software architecture and data acquisition pipeline
were reliable (they were), and (3) establishing two experimental baselines for
the Freeplay Sandbox task: the `human' baseline on one hand (child-child
condition), an `asocial' baseline on the other hand (child - \emph{non-social}
robot condition). These two baselines are situated at the two ends of the
spectrum of social interaction. They aim at characterising the qualitative and
quantitative bounds of this social spectrum and can be used by the research
community to evaluate given interaction policies.

A detailed description of the dataset is outside of the scope of this paper, and
we only provide hereafter cursory informations on the dataset. Specific details
regarding the methodology and the acquisition procedure can be found on the
dataset website: [hidden for blind review]. The dataset is open and accessible
to any interested researcher.

In total, 120 children were recorded for a total duration of 45 hours and 48
minutes of data collection. These 120 children (age 4 to 8) were split into two
conditions: a child-child condition and a child-robot condition. In both
condition, and after a short tutorial, the children were simply invited to
freely play with the sandbox, for as long as they wished (with a cap at 40 min).

In the child-child condition (as seen in Figure~\ref{annotator}), 45 freeplay interactions (\ie 90 children) were
recorded with a duration M=24.15 min (SD=11.25 min).

In the child-robot condition, 30 children were recorded, M=19.18 min (SD=10
min). In this later condition, the robot
behaviour was coded to be purposefully \emph{asocial}: the robot would autonomously play with
the game items, but would avoid any social interaction (no social gaze, no
verbal interaction, no reaction to the child-initiated game actions).

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{annotator}
    \caption{Screenshot of the dedicated tool developed for rapid annotation of
    the social interactions.}
    \label{annotator}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Discussion}
\label{sec:discussion}

\subsection{Analysis of the free play sandbox}

Our free play sandbox elicits a loosely structured play situation. The
situation is indeed essentially aimless (free play). The goals and play episodes
that we observe during the interactions are not pre-defined, and essentially
unknown until they are decided by the players. In this sense, the free play
sandbox permits and elicits what is fundamentally \emph{unstructured play}.
\todo{Is it really aim-less? Children create their own goals, no?}

However, a second look reveals some important structuring elements.

First, the physical bounds of the sandbox (an interactive table) limit the
play zone to a well defined and relatively small area. As a consequence,
children are mostly static (they are sitting in front of the table) and their
primary form of interaction is 2D pick and place of items by drag and dropping
them on the screen.

Second, the game items themselves structure the game scenarios. Items are either
iconic characters (animals or children) or plain blocks
(Figure~\ref{fig|sandbox}). The former have strong semantics associated to them
(like 'crocodiles like water and eat children'), while the later have little
semantics associated. The game background, with its recognizable zones, also
elicit a particular type of games (like building a zoo or pretending we explore
the savannah).

Two more elements (the role and place of the experimenter, and the initial
prompt given to the participant) have an important impact on the shape of the
play situation. We discuss separately these two points in the description of the
experimental procedure.

Overall, the free play sandbox supports a \emph{loosely structured} form of play: the
actual play situations are not known and might change several times during the
interaction; the game actions, even though all based on a single modality (picking and
placing game items), are unlimited; the social interactions between participants
are multi-modal (speech, body postures, gestures, facial expressions, etc.) and
unconstrainted. However, the broad domain of the play situations and the range of
possible social interactions is bound by the physical bounds of the play zone
and the theme of the game items.
\todo{more than one actions, also drawing}

\paragraph{Forms of play elicited by the sandbox}

\begin{itemize}
    \item dramatic play
    \item storytelling
    \item creative play
    \item pretend play
\end{itemize}

\subsection{An exploration of the dynamics of social interactions}

\begin{figure}
\centering

\resizebox{0.8\linewidth}{!}{%
\begin{tikzpicture}[
                font=\sffamily,
                >=latex,
                every edge/.style={draw, very thick},
                node/.style={draw, rounded corners, align=center, inner sep=5pt, fill=black!20},
                label/.style={midway, align=center, font=\scriptsize\sffamily, fill=white}]

    %%% NODES
    \node [node] (mentalrepresentations) {Mental Representations};
    \node [above left=of mentalrepresentations] (gameactions) {\it game actions};
    \node [node, above=of gameactions] (gamedynamics) {Game dynamics};
    \node [above right=of mentalrepresentations] (socialactions) {\it social actions};
    \node [node, above=of socialactions ] (socialdynamics) {Social Dynamics};

    %%% CONNECTIONS
    \path (gamedynamics) edge [<->, dashed] node[label, above] {?}(socialdynamics);
    \path (gamedynamics) edge [<-, bend right] node[label] {mediate}(gameactions);
    \path (socialactions) edge [->, bend left] node[label] {reflect}(mentalrepresentations);
    \path (socialactions) edge [->, dashed] node[label,above] {?}(gamedynamics);
    \path (gameactions) edge [->, dashed] node[label,above] {?}(socialdynamics);
    \path (gameactions) edge [->, bend right] node[label] {reflect}(mentalrepresentations);
    \path (socialdynamics) edge [<-, bend left] node[label] {shape}(socialactions);

\end{tikzpicture}
}
\label{fig|researchquestions}
\caption{Game dynamics, social dynamics and mental representations are linked
    with each other through social actions (implicit and explicit communication)
    and actions on the environment (mainly game actions)}
\end{figure}

The mechanisms underlying the social interactions happening during free play
raise important questions for social robotics.

\paragraph{Game Dynamics}

\begin{itemize}
    \item do we observe a sequence of sub-games?
    \item how are these sub-games initiated?
    \item how does A keep track of what B is playing at?
    \item how do we segment the action flow into meaningful play situations?
\end{itemize}

\paragraph{Social Dynamics}

\begin{itemize}
    \item does a ``social protocol'' establish?
    \item if so, what are these social rules?
    \item implicit vs explicit rule setting?
    \item what interaction modalities (or combination thereof) are relied upon
        to establish and maintain this ``social protocol''?
\end{itemize}

\paragraph{Mental representations}

\begin{itemize}
    \item Does A know what B is thinking of at time t?
    \item Do the participants perform explicit grounding of their respective mental models
        (ie, ask the other one what she is doing/thinking/planning to do?)
    \item How does the implicit grounding take place? attention tracking?
    \item Do we need at all to know what the other intends to do to actually
        collaborate? surface alignment vs deep grounding
    \item Procedural (how do you do it?) vs Semantic
        grounding (what are you doing? playing at?)
\end{itemize}

\subsection{Stages of Play}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion}
\label{sec:conclusion}

\section*{Acknowledgments}

This work has been supported by the EU H2020 Marie Sklodowska-Curie Actions
project DoRoThy (grant 657227).

\bibliographystyle{ACM-Reference-Format}
\bibliography{biblio}
\end{document}

### 2017-10-05 paper.tex
Update on Overleaf.

@@
 their most unconstrained version (like high dimensionality and uncertainties),
-we have developed a \emph{free play sandbox}, Figure~\ref{fig|setup}: a large
+we have developed a \emph{free play sandbox}, Figure~\ref{fig|freeplay}: a large
 touchscreen (60cm $\times$ 33cm) is used as an interactive surface
@@
 
-observing and quantifying social interactions.
+The Freeplay Sandbox provides the opportunity to observe children interacting in a natural way in an open but framed setup.
+As the system can run on a single computer platform it can easily be deployed in the 'wild', in places where the children naturally interact such as classroom.
+The quantity and thoughtfulness of information logged allows to keep a track of every interaction happening around the game.
+
+These advantages combined with the openness of the task proposed make this setup a powerful tool to observe and quantify a large spectrum of social behaviours expressed by children when interacting in a natural environment (might be interesting to add a list here).
+The compactness of the system makes it easy to compare data from different locations.
 
@@
 This setup allows to explore the impact of different social behaviours on the
-children in an open yet standardised environment.
+children independently of the 'game policy' controlling by the robot.
 
@@
 
-Standardised method to gather data and use deep learning for automated
-annotation.
+With the quantity of data logged and the high number of interaction achievable with the Freeplay Sandbox, it supports the type of requirement for recent Machine Learning approaches such as deep learning.
+The similar position of the children in all interactions makes the combination of data from different interaction easier than other less compact systems.
 
-Can be used to explore the impact of social behaviours of the robot: "game
-action policy" can be identical, but more social behaviours can be desired
-
-Can be used to explore learning of social action policy for a robot.
+From the information collected on the children, social behaviours can be extracted and used on a robot.
 
@@
 to any interested researcher.
+\todo{Probably add "an anonymised version of the dataser is open..." Reviewers might be concerned about ethics}
 

### 2017-10-05 paper.tex
Reworked the abstract

@@
 with constrained protocols to allow for robust measurements and a degree of
-replicability, but at the cost of ecological validity. Or on the contrary,
-studies are run \emph{in the wild}, which may provide better experimental
-realism, but often at the expense of rigorous interaction metrics and replicability.
+replicability, but at the cost of ecological validity; or \emph{in the wild},
+which leads to superior experimental realism, but often with limited
+replicability and at the expense of rigorous interaction metrics.
 
-We introduce in this paper a novel interaction situation designed for ecological
-validity while ensuring `good' scientific properties (replicable, clear metrics,
-possibility of both autonomous and Wizard-of-Oz robot behaviours).
-This task focuses on child-robot interaction in a \emph{sandboxed} free-play
-environment. We present hereafter the rationale for this task, its technical and
-practical aspects (including the open-source source code), as well as two
-large datasets that represent interaction baselines for this task.
+We introduce in this paper a novel interaction paradigm, designed to elicit rich
+and varied social interactions while exhibiting `good' scientific properties
+(replicable, clear metrics, possibility of either autonomous or Wizard-of-Oz
+robot behaviours). This paradigm focuses on child-robot interactions, and
+builds on a sandboxed free-play environment. We present hereafter the
+rationale and design of this interaction paradigm, its associated
+methodological and technical aspects (including the open-source
+implementation of the software platform), as well as two large open datasets
+acquired with this paradigm, and meant to act as experimental baselines for
+future research.
 

### 2017-10-05 paper.tex
Re-structured first part of intro

@@
 
-Evaluating in a rigorous manner human-robot social interactions is notoriously
-difficult~\cite{baxter2016characterising}: studies are either conducted in labs
-with constrained protocols to allow for robust measurements and a degree of
-replicability, but at the cost of ecological validity. Or on the contrary,
-studies are run \emph{in the wild}, which may provide better experimental
-realism, but often at the expense of rigorous interaction metrics and replicability.
+\subsection{Studying social interactions}
+
+Studying social interactions requires a social \emph{situation} that
+effectively elicits interactions between the participants. Such a situation is
+typically scaffolded by a social task, and consequently, the nature of this task
+influences in fundamental ways the kind of interactions that might be observed
+and analysed. In particular, the socio-cognitive tasks commonly found in the
+literature of experimental psychology (and HRI) often have a narrow focus:
+because they aim at studying one (or a few) specific social or cognitive skill
+in isolation and in a controlled manner, these tasks are typically simple and
+highly constrained (for instance, an object hand-over task; a perspective-taking
+task with cubes, etc.). While these focused endeavours are important and
+necessary, we -- as a community -- also acknowledge that these interaction
+scenarios do not reflect the complexity and dynamics of real-world
+interactions~\cite{baxter2016characterising}, and we certainly observe a strong
+trend within our community towards capturing, interpreting and acting upon the rich set of
+naturally-occurring social interactions.
+
+Specifically, we believe that further progress in the study of human-robot
+interactions should be scaffolded by socio-cognitive challenges that:
+
+\begin{itemize}
+    \item are long enough and varied enough to elicit a large range of
+        interaction situations;
+    \item foster rich multi-modal interaction, such as simultaneous speech, gesture, and gaze
+        behaviours;
+    \item are loosely directed, to maximise natural, non-contrived behaviours;
+    \item evidence complex social dynamics, such as rhythmic coupling, joint attention,
+        implicit turn-taking;
+    \item include a certain level of non-determinism and unpredictability.
+\end{itemize}
+
+The challenge lies in designing a social task that exhibits these features
+\emph{while maintaining `good' scientific properties} (repeatability, replicability,
+robust metrics) as well as good practical properties (not requiring unique or
+otherwise very costly experimental environments, not requiring very specific
+hardware or robotic platform, easy deployment, short enough experimental
+sessions to allow for large groups of participants).
+
+\subsection{Social play}
 
@@
 scheme for social interaction).
-
-\subsection{Studying social interactions}
-
-Studying social interactions requires a social situation that elicits
-interactions between the participants.  Such situation is typically scaffolded
-by a social task, and consequently, the nature of this task influences in
-fundamental ways the kind of interactions that might be observed and analysed.
-
-Because they typically target specific social or cognitive skills in a tightly
-controlled manner, traditional socio-cognitive tasks are usually simple,
-constrained tasks that do not reflect the complexity and dynamics of real-world
-interactions. Conversely, state-of-the-art human-robot interaction aims at
-capturing, interpreting and acting upon a rich set of
-naturally-occurring social interactions. As such, we argue that further advance
-in social human-robot interaction requires us to study and tackle tasks that:
-
-\begin{itemize}
-    \item are loosely directed, to maximise natural, non-contrived behaviours;
-    \item are long enough and varied enough to elicit a large range of
-        interaction situations;
-    \item foster rich multi-modal interaction, such as simultaneous speech, gesture, and gaze
-        behaviours;
-    \item evidence complex social dynamics, such as rhythmic coupling, joint attention,
-        implicit turn-taking, a level of unpredictability.
-\end{itemize}
-
-The challenge lies in designing a social task that exhibits these features,
-while maintaining `good' scientific properties (repeatability, replicability,
-robust metrics) as well as good practical properties (not requiring unique or
-otherwise very costly experimental environments, not requiring very specific
-hardware or robotic platform, easy deployment, short enough experimental
-sessions to allow for large groups of participants). So far, our community has
-struggled with this endeavour~\cite{baxter2016characterising}.
 

### 2017-10-05 paper.tex
Reworked the end of the intro and the begining of section 2

@@
 
+We introduce in this paper such a task, designed to elicit rich, complex, varied
+social interactions while being well suited for interactions with robots and
+supporting rigorous scientific methodologies.
+
 \subsection{Social play}
 
-Existing benchmarks (robocup at home) have issues -- the next challenge for HRI
-is to tackle complex, unpredictable situations -> free play
+Our interaction paradigm is based on free and playful interactions (free play)
+in a \emph{sandboxed} environment: while the interaction is free (participants
+are not directed to perform any particular task beyond playing), the activity is
+both \emph{scaffolded} and \emph{constrained} by the setup mediating the
+interaction (essentially, a large table-top touchscreen).
+Participant engage in open-ended and non-directive play situations, yet sufficiently
+well defined to be reproducible and practical to record and analyse.
 
-This task is based on free child-robot interactions (free play) in a
-\emph{sandboxed} environment: while the interaction is free (children do not
-have a particular task to perform, except to play), the activity is scaffolded
-by a specific apparatus (a large touchscreen) where the interaction take place.
-Capturing the entirety of the interaction is consequently possible and
-practical. We present in the remaining of the paper the methodology and
-experimental procedure associated to the free-play task, as well as the main
-analyses that can be conducted on the collected data (including a new annotation
-scheme for social interaction).
-
-children can engage in open-ended and
-non-directive play situations, yet sufficiently
-defined to be reproducible;
-
-\subsection{Background on social play}
+This initial description frames the socio-cognitive interactions that might be
+observed and studied: playful, dyadic, face-to-face interactions. While gestures
+and manipulations (including joint manipulations) play an important role in this
+paradigm, the participants do not typically move much during the interaction.
+Because it builds on play, this paradigm is also naturally suited to the study
+of child-child and child-robot interactions.
 
@@
 in that field builds on the precursory work of Parten who established five
-\emph{stages of play}~\cite{parten1932social}:
+\emph{stages of play}~\cite{parten1932social}, corresponding to different stages
+of development, and accordingly associated to typical age ranges:
 
 \begin{enumerate}
-    \item {\bf Solitary (independent) play}: Playing separately from
+    \item {\bf Solitary (independent) play}, age 2-3: Playing separately from
         others, with no reference to what others are doing.
-    \item {\bf Onlooker play}: Watching others play. May engage in
+    \item {\bf Onlooker play}, age 2.5-3.5: Watching others play. May engage in
         conversation but not engaged in doing. True focus on the children at
         play.
-    \item {\bf Parallel play} (adjacent play, social co-action): Playing
+    \item {\bf Parallel play} (adjacent play, social co-action), age 2.5-3.5: Playing
         with similar objects, clearly beside others but not with them (near
         but not with others.)
-    \item {\bf Associative play}:  Playing with others without
+    \item {\bf Associative play}, age 3-4:  Playing with others without
         organization of play activity. Initiating or responding to
         interaction with peers. 
-    \item {\bf Cooperative play}: Coordinating one's behavior with that
+    \item {\bf Cooperative play}, age 4+: Coordinating one's behavior with that
         of a peer. Everyone has a role, with the emergence of a sense of
@@
 
-\todo{Should the age be added?}
+These five stages of play have been extensively discussed and refined over the
+last century, yet remain remarkably widely accepted as such. It must be noted
+that the age ranges are only indicative. In particular, most of the early
+behaviours still occur at times by older children.
 
-These five stages of play have been extensively discussed and refined over the
-last century, yet remain remarkably widely accepted as such, and to the ears of
-researchers in social robotics, they may resonate as an implicit roadmap.
-
-\subsection{Framing: face-to-face dyadic and sandboxed child-robot interactions}
-
-To mitigate the technical challenges presented by free play interactions in
-their most unconstrained version (like high dimensionality and uncertainties),
-we have developed a \emph{free play sandbox}, Figure~\ref{fig|freeplay}: a large
-touchscreen (60cm $\times$ 33cm) is used as an interactive surface
-(\emph{sandtray} paradigm~\cite{baxter2012touchscreen}). Two agents (either a
-pair of children or one child and one robot) play together by freely moving on a
-2D plane items (characters and blocks, see Figure~\ref{fig|sandbox}). A
-background image depicts a generic empty environment, with different symbolic
-zones (water, grass, path, bushes...). Additionally, the children can draw on
-the background using a tool applying a color on their touches. The players do
-not have any particular task to complete, they are simply invited to freely
-play.
+Interestingly, these five stages can been looked at from the perspective of HRI
+as well. They certainly evoke a roadmap for the development of human-robot
+social interactions.
 
@@
 \subsection{Task}
-description of the task and why it fills a gap in the literature/tools used so
-far (provide clearer,  and interesting setup, \emph{easy} to reuse)
+
+To mitigate the technical challenges presented by free play interactions in
+their most unconstrained version (like high dimensionality and uncertainties),
+we have developed a \emph{free-play sandbox}, Figure~\ref{fig|freeplay}: a large
+touchscreen (60cm $\times$ 33cm) is used as an interactive surface
+(\emph{sandtray} paradigm~\cite{baxter2012touchscreen}). Two participants (either a
+pair of children or one child and one robot) play together by freely moving on a
+2D plane items (characters and blocks, see Figure~\ref{fig|sandbox}). A
+background image depicts a generic empty environment, with different symbolic
+zones (water, grass, path, bushes...). Additionally, the children can draw on
+the background using a tool applying a color on their touches. The players do
+not have any particular task to complete, they are simply invited to freely
+play. Importantly, they can play for as long as they wish (for
+practical reasons, we have limited the sessions to a maximum of 40 minutes in
+our own experiments, see Section~\ref{sec:dataset}).
+
+Capturing all the interactions taking place during the play sessions is possible
+and practical with this setup. Even though the children will typically move a
+little, the task is fundamentally a face-to-face, spatially delimited,
+interaction, and as such simplifies the data collection. For instance, during
+our dataset acquisition campaign (120 children, more than 45h of footage), the
+children's faces were automatically detected in 97\% [TODO: check that!] of the
+recorded frames.
 
@@
 
+TODO: percentage faces seen?
+
 \begin{figure}

### 2017-10-06 paper.tex
Update on Overleaf.

@@
 
-Evaluating in a rigorous manner human-robot social interactions is notoriously
-difficult~\cite{baxter2016characterising}: studies are either conducted in labs
+Evaluating human-robot social interactions in a rigorous manner is notoriously
+difficult: studies are either conducted in labs
 with constrained protocols to allow for robust measurements and a degree of
@@
 
-We introduce in this paper a novel interaction paradigm, designed to elicit rich
-and varied social interactions while exhibiting `good' scientific properties
-(replicable, clear metrics, possibility of either autonomous or Wizard-of-Oz
+We introduce a novel interaction paradigm, designed to elicit rich
+and varied social interactions while having desirable scientific properties
+(replicability, clear metrics, possibility of either autonomous or Wizard-of-Oz
 robot behaviours). This paradigm focuses on child-robot interactions, and
-builds on a sandboxed free-play environment. We present hereafter the
-rationale and design of this interaction paradigm, its associated
+builds on a sandboxed free-play environment. We present the
+rationale and design of the interaction paradigm, its 
 methodological and technical aspects (including the open-source

### 2017-10-06 paper.tex
Update on Overleaf.

@@
 
-\section{The difficult evaluation of social interactions}
+\section{The challenges in evaluating social interactions}
 \label{sec:intro}
@@
 literature of experimental psychology (and HRI) often have a narrow focus:
-because they aim at studying one (or a few) specific social or cognitive skill
+because they aim at studying one (or a few) specific social or cognitive skills
 in isolation and in a controlled manner, these tasks are typically simple and

### 2017-10-06 paper.tex
Update on Overleaf.

@@
 
-We introduce in this paper such a task, designed to elicit rich, complex, varied
+In this paper, we introduce such a task, designed to elicit rich, complex, varied
 social interactions while being well suited for interactions with robots and
@@
 situations and social behaviours that \emph{play} elicits. Most of the research
-in that field builds on the precursory work of Parten who established five
+in this field builds on the early work of Parten who established five
 \emph{stages of play}~\cite{parten1932social}, corresponding to different stages
-of development, and accordingly associated to typical age ranges:
+of development, and accordingly associated with typical age ranges:
 
@@
     \item {\bf Onlooker play}, age 2.5-3.5: Watching others play. May engage in
-        conversation but not engaged in doing. True focus on the children at
+        conversation but not engage in doing. True focus on the children at
         play.
@@
 
-To mitigate the technical challenges presented by free play interactions in
-their most unconstrained version (like high dimensionality and uncertainties),
-we have developed a \emph{free-play sandbox}, Figure~\ref{fig|freeplay}: a large
-touchscreen (60cm $\times$ 33cm) is used as an interactive surface
-(\emph{sandtray} paradigm~\cite{baxter2012touchscreen}). Two participants (either a
-pair of children or one child and one robot) play together by freely moving on a
-2D plane items (characters and blocks, see Figure~\ref{fig|sandbox}). A
-background image depicts a generic empty environment, with different symbolic
-zones (water, grass, path, bushes...). Additionally, the children can draw on
-the background using a tool applying a color on their touches. The players do
+We have designed a new experimental task, called the \emph{free-play sandbox},
+that is based on free play interactions. Pairs of children (4-8 years old) are
+invited to freely draw and interact with items displayed on an interactive
+table, without any explicit goal set by the experimenter
+(Fig.~\ref{fig|freeplay}).  The task is designed so that children can engage in
+open-ended and non-directive play, yet it is sufficiently constrained to be
+suitable for recording, and allows the reproduction of social behaviour by an
+artificial agent in comparable conditions.
+
+The background image depicts a contextual environment, with different 
+zones (for example, water, grass, path, bushes...). Additionally, the children can draw on
+the background using a tool of the colour is costumisable. The players do
 not have any particular task to complete, they are simply invited to freely
@@
 our dataset acquisition campaign (120 children, more than 45h of footage), the
-children's faces were automatically detected in 97\% [TODO: check that!] of the
+children's faces were automatically detected in 97\% \textcolor{red}{[TODO: check that!]} of the
 recorded frames.
@@
 
-We have designed a new experimental task, called the \emph{free-play sandbox},
-that is based on free play interactions. Pairs of children (4-8 years old) are
-invited to freely draw and interact with items displayed on an interactive
-table, without any explicit goal set by the experimenter
-(Fig.~\ref{fig|freeplay}).  The task is designed so that children can engage in
-open-ended and non-directive play, yet it is sufficiently constrained to be
-suitable for recording, and allows the reproduction of social behaviour by an
-artificial agent in comparable conditions.
-
 \begin{figure}[ht!]
@@
 
-The free-play sandbox follows the sandtray
-paradigm~\cite{baxter2012touchscreen}: a large touchscreen (60cm $\times$ 33cm,
-with multitouch support) is used as an interactive surface (\emph{sandtray}).
-Two children play together by freely moving interactive items
-(Fig.~\ref{fig|sandbox}) on the surface. A background image depicts a generic
-empty environment, with different symbolic colours (water, grass, beach,
-bushes...). By drawing on top of the background picture, the children can change
-the environment to their liking. The players do not have any particular task to
-complete, they are simply invited to freely play for as long as they want.
+% TB: ALSO EXPLAINED ABOVE, THIS PARAGRAPH IS REDUNDANT
+%
+% The free-play sandbox follows the sandtray
+% paradigm~\cite{baxter2012touchscreen}: a large touchscreen (60cm $\times$ 33cm,
+% with multitouch support) is used as an interactive surface (\emph{sandtray}).
+% Two children play together by freely moving interactive items
+% (Fig.~\ref{fig|sandbox}) on the surface. A background image depicts a generic
+% empty environment, with different symbolic colours (water, grass, beach,
+% bushes...). By drawing on top of the background picture, the children can change
+% the environment to their liking. The players do not have any particular task to
+% complete, they are simply invited to freely play for as long as they want.
 
-\todo{almost identical to 1.3 and mention dataset}
-
-In order to constrain the dataset to a tractable domain, the task is limited to
-a dyadic face-to-face interaction.  The resulting interaction space is
-nevertheless high-dimensional, with two main foci: the game-related actions
-(drawing, manipulating items, telling stories) and the social interactions
-(monitoring the other's actions, discussing and agreeing on joint actions,
-supportive/antagonistic behaviours, etc.).
+% TB: ALSO EXPLAINED ABOVE, THIS PARAGRAPH IS REDUNDANT
+%
+% \todo{almost identical to 1.3 and mention dataset}
+% In order to constrain the dataset to a tractable domain, the task is limited to
+% a dyadic face-to-face interaction.  The resulting interaction space is
+% nevertheless high-dimensional, with two main foci: the game-related actions
+% (drawing, manipulating items, telling stories) and the social interactions
+% (monitoring the other's actions, discussing and agreeing on joint actions,
+% supportive/antagonistic behaviours, etc.).
 
@@
 
+\todo{Very much unfinished section.}
 focus on abstract socio-cognitive facets (robot
@@
 This setup allows to explore the impact of different social behaviours on the
-children independently of the 'game policy' controlling by the robot.
+children independently of the `game policy' controlling by the robot.
 
@@
 aspect is of a lower importance, the robot can be controlled by a human rather
-than an algorithm. This paradigm, where a human teleoperate a robot to interact
+than an algorithm. This paradigm, where the robot is tele-operated to interact
 with a naive partner is called Wizard of Oz (WoZ) and is used in numerous
@@
     \caption{Software architecture of the free-play sandbox. Left (purple) nodes
-    are tided to the sandtray (game interface (1) and camera drivers (2)). Nodes in the
+    are connected to the sandtray (game interface (1) and camera drivers (2)). Nodes in the
     centre (green) implement the behaviour of the robot (play policy (3) and robot
@@
     are published as well, and turned into an occupancy map, used to plan the
-    robot's arm motion.
-    visualisation tool.}
+    robot's arm motion.}
     \label{fig|rviz}
@@
 \subsection{Interactive game}
-The interactive game is coded using QML, and possess main background image on
-top of which items (animals, humans and objects) can be moved from one part of
-the world to another. The children can also use a drawing mode in which some
-create strokes of color on a layer between the background and the items. This
-increase the openess of the game as symbolic objects can be added by drawing
+The interactive game is coded using QML, and displays a main background image on
+top of which items (animals, humans and objects) can be moved. 
+The children can also use a drawing mode to
+create coloured strokes on a layer between the background and the items. This
+increases the openess of the game as symbolic objects can be added by drawing
 (eg. Figure~\ref{fig|sandbox}. The game exposes the image of the background
@@
 regarding the methodology and the acquisition procedure can be found on the
-dataset website: [hidden for blind review]. The dataset is open and accessible
+dataset website: \url{https://freeplay-sandbox.github.io/}. The dataset is open and accessible
 to any interested researcher.
@@
 
-This work has been supported by the EU H2020 Marie Sklodowska-Curie Actions
-project DoRoThy (grant 657227).
+Redacted for anonymous review.
+%This work has been supported by the EU H2020 Marie Sklodowska-Curie Actions
+%project DoRoThy (grant 657227).
 

### 2017-10-06 paper.tex
Update on Overleaf.

@@
 The children can also use a drawing mode to
-create coloured strokes on a layer between the background and the items. This
-increases the openess of the game as symbolic objects can be added by drawing
-(eg. Figure~\ref{fig|sandbox}. The game exposes the image of the background
-with the drawing in case of change and the positions of each item as a TF frame.
+create coloured strokes on a layer between the background and the items, which adds another layer of unconstrained interaction to the game
+(eg. Figure~\ref{fig|sandbox}. The game exposes the image of the background,
+the drawings, and the positions of the objects as a TF frame.
 
 \subsection{Sensors}
-Two RGB-D cameras have a fixed position on the Sandtray frame and face one child
-each. This fixed position of the cameras permit to extract social features from
-the images such as head gaze, gaze, pointing and position in a general
-coordinate system usable by each node. In addition of the image, the cameras
-posses a microphone to listen and record what the children are saying.
+Two RGB-D cameras are mounted at fixed positions on the Sandtray frame, each facing a child. This fixed position of the cameras allows the extraction of social features 
+such as head pose, eye gaze, pointing and body posture, all described in a general
+coordinate system available to each software node. In addition to the image, the cameras
+posses a microphone to record the auditory aspect of the interaction.
 
 \subsection{Supervisor}
-A web interface can be used by an experimenter to manage the meta parameters of
-the interaction: demographics, start-stop, task selected, recording...
+A web interface can be used by the experimenter to manage the meta parameters of
+the interaction: log data on the subjects, start-stop, select a task, manage data recording...
 
 \subsection{Robot Control}
-As stated in section~\ref{ssec|CRI}, a robot can interact in place of one of the
+As stated in section~\ref{ssec|CRI}, a robot can act as play partner instead of one of the
 children. This robot can either be autonomous selecting actions based on the

### 2017-10-06 paper.tex
Checking with Emmanuel what remains to be done

@@
 \maketitle
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\columnwidth]{setup_top}
+    %\hspace{1em}
+    %\includegraphics[width=0.45\columnwidth]{setup_child_robot_top}
+    %\includegraphics[width=0.45\columnwidth]{child-child-env}
+    \caption{The free-play social interactions sandbox: two children interact in
+    a free-play situation, by drawing and manipulating items on a touchscreen.
+    Children are facing each other and sit on cushions. Each child wears a
+    bright sports bib, either purple or yellow, to facilitate later
+    identification.}
+
+    \label{fig|freeplay}
+\end{figure}
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.9\columnwidth]{setup_top}
-    %\hspace{1em}
-    %\includegraphics[width=0.45\columnwidth]{setup_child_robot_top}
-    %\includegraphics[width=0.45\columnwidth]{child-child-env}
-    \caption{The free-play social interactions sandbox: two children interact in
-    a free-play situation, by drawing and manipulating items on a touchscreen.
-    Children are facing each other and sit on cushions. Each child wears a
-    bright sports bib, either purple or yellow, to facilitate later
-    identification.}
-
-    \label{fig|freeplay}
-\end{figure}
-
 \begin{figure}[ht!]
@@
 
-\section{Technical Implementation}
+\section{Task Implementation}
 \label{sec:impl}
@@
 
-\subsection{Supervisor}
-A web interface can be used by the experimenter to manage the meta parameters of
-the interaction: log data on the subjects, start-stop, select a task, manage data recording...
+\todo{briefly present ROS pipeline for sensors}
 
@@
 
-\subsection{Records and logs}
+\subsection{Experiment Manager}
 
-Maybe redundant with Data  collection.
-
-\subsection{Tools}
+A web interface can be used by the experimenter to manage the meta parameters of
+the interaction: log data on the subjects, start-stop, select a task, manage data recording...
 
@@
 
+\subsection{Protocol}
+
+\todo{TBD}
+
 \subsection{Data collection}
@@
 
-\subsection{RGB-D streams analysis}
+\subsection{Data processing}
 
@@
 
-\subsection{Audio processing}
+\paragraph{Audio processing}
 
@@
 
-\subsection{Game interactions analysis}
+\paragraph{Game interactions analysis}
+
+\todo{check section title}
 
@@
 \subsection{Video Coding}
+
+\todo{Mention the annotator tool}
 

### 2017-10-06 paper.tex
Added plot of duration of interactions
Added the data acquisition protocol

@@
 to pre-process it, and analyse it. We call them \emph{canonical} because they
-are either well integrated into the software pipeline of the sandbox (\eg ROS
-integration), represent start-of-the-art techniques, or result from the 
-synthesis of existing techniques (this is especially true of the coding scheme
-that we introduce for the annotation of social behaviours,
-Section~\ref{sec|coding-scheme} below).
+are standard procedures, and where relevant, well integrated into the software pipeline of the sandbox (\eg ROS
+integration) and represent state-of-the-art techniques. For the specific purpose
+of manually annotating the social interaction, we introduce as well a novel
+coding scheme, resulting from the synthesis of several existing techniques
+(Section~\ref{sec|coding-scheme} below).
 
-However, they do not represent normative procedures. Researchers interested in
+However, these procedure are not normative. Researchers interested in
 reusing the free-play sandbox task for their own research would naturally adapt
@@
 
-\todo{TBD}
+We typically adhere to the acquisition procedure described in
+Table~\ref{tab|protocal} with all participants.  To ease later identification,
+each child is also given a different and brightly coloured sports bib to wear.
+
+Importantly, during the \emph{Greetings} stage, we show the robot both moving and
+speaking (for instance, ``Hello, I'm Nao. Today I'll be playing with you.
+Exciting!'' while waving at the children). This is meant to set the children's
+expectations: they have seen that the robot can speak, move, and even behave in
+a social way.
+
+Also, the game interface of the free-play sandbox offers a tutorial mode, used
+to ensure the children know how to manipulate items on a touchscreen and draw.
+In our experience, this has never been an issue for children.
+
+\newcommand{\tabitem}{~~\llap{\textbullet}~~}
+
+\begin{table}[!]
+\caption{Data acquisition protocol}
+    \label{tab|protocol}
+\centering
+\begin{tabular}{p{\linewidth}}
+\toprule
+\bf Greetings \emph{(about 5 min)} \\
+\tabitem explain the purpose of the study: showing robots how children play  \\
+\tabitem briefly present a Nao robot: the robot stands up, gives a short
+message, and sits down. \\
+\tabitem place children on cushions \\ 
+\tabitem complete demographics on the tablet \\
+\tabitem remind the children that they can withdraw at anytime \\ \midrule
+\bf Tutorial \emph{(1-2 min)}  \\
+explain how to interact with the game, ensure the children are confident
+with the manipulation/drawing \\ \midrule
+\bf Freeplay task (up to 40 min)\\
+\tabitem initial prompt: \emph{"Just to remind you, you can use the animals or draw. Whatever you
+like. If you run out of ideas, there's also an ideas box. For example, the first one is a
+zoo. You could draw a zoo or tell a story. When you get bored or don't want to play
+anymore, just let me know."} \\
+\tabitem let children play \\
+\tabitem once they wish to stop, stop recording \\ \midrule
+\bf Debriefing      \emph{(about 2 min)} \\
+\tabitem answer possible questions from the children \\
+\tabitem give small reward (\eg stickers) as a thank you \\ \bottomrule
+\end{tabular}
+\end{table}
 
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\columnwidth]{durations}
+    %\hspace{1em}
+    %\includegraphics[width=0.45\columnwidth]{setup_child_robot_top}
+    %\includegraphics[width=0.45\columnwidth]{child-child-env}
+    \caption{Durations of the interactions for the two conditions.}
+
+    \label{fig|durations}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-10-06 paper.tex
Add reference to figure in implementation

@@
 
-\subsection{Interactive game}
-The interactive game is coded using QML, and displays a main background image on
-top of which items (animals, humans and objects) can be moved. 
-The children can also use a drawing mode to
-create coloured strokes on a layer between the background and the items, which adds another layer of unconstrained interaction to the game
-(eg. Figure~\ref{fig|sandbox}. The game exposes the image of the background,
-the drawings, and the positions of the objects as a TF frame.
+\subsection{Interactive game} The interactive game
+(Fig.~\ref{fig|architecture}.1) is coded using QML, and displays a main
+background image on top of which items (animals, humans and objects) can be
+moved.  The children can also use a drawing mode to create coloured strokes on a
+layer between the background and the items, which adds another layer of
+unconstrained interaction to the game (eg. Figure~\ref{fig|sandbox}. The game
+exposes the image of the background, the drawings, and the positions of the
+objects as a TF frame.
 
-\subsection{Sensors}
-Two RGB-D cameras are mounted at fixed positions on the Sandtray frame, each facing a child. This fixed position of the cameras allows the extraction of social features 
-such as head pose, eye gaze, pointing and body posture, all described in a general
-coordinate system available to each software node. In addition to the image, the cameras
-posses a microphone to record the auditory aspect of the interaction.
+\subsection{Sensors} Two RGB-D cameras are mounted at fixed positions on the
+Sandtray frame, each facing a child. This fixed position of the cameras allows
+the extraction of social features such as head pose, eye gaze, pointing and body
+posture, all described in a general coordinate system available to each software
+node. In addition to the image, the cameras posses a microphone to record the
+auditory aspect of the interaction. One node per camera
+(Fig.~\ref{fig|architecture}.2) outputs on dedicated topics the audio and video
+streams.
 
@@
 
-\subsection{Robot Control}
-As stated in section~\ref{ssec|CRI}, a robot can act as play partner instead of one of the
+\subsection{Robot Control} As stated in section~\ref{ssec|CRI}, a robot
+(Fig.~\ref{fig|architecture}.9) can act as play partner instead of one of the
 children. This robot can either be autonomous selecting actions based on the
@@
 controller. The position of every item is exposed as a TF frame, the background
-is segmented in zones of identical colors, social element of the state the
-interaction are collected through the RGBD camera and the microphone facing the
-child. As visible on Figures~\ref{fig|freeplay} and \ref{fig|rviz}, the camera
-covers the head of the child as well as most of the upperbody, and applying
-libraries such as DLib and OpenPose, the position of facial feature and skeleton
-of the child are extracted and can be used to obtain: head gaze, gaze and
-gestures such as pointing. All these inputs can be combined to provide the robot
-with more social inputs to test the sociability of a robotic controller and its
-impact on the interaction.
+is segmented in zones of identical colors (Fig.~\ref{fig|architecture}.5),
+social element of the state the interaction are collected through the RGBD
+camera and the microphone facing the child. As visible on
+Figures~\ref{fig|freeplay} and~\ref{fig|rviz}, the camera covers the head of the
+child as well as most of the upperbody, and applying libraries such as DLib and
+OpenPose, the position of facial feature and skeleton of the child are extracted
+and can be used to obtain: head gaze, gaze and gestures such as pointing. All
+these inputs can be combined to provide the robot with more social inputs to
+test the sociability of a robotic controller (Fig.~\ref{fig|architecture}.3) and
+its impact on the interaction.
 
@@
 touchscreen before the start of the interaction, so the transformation between
-the robot coordinate system and the touchscreen is known. And this robot
-location can also be used to identify gazes from the child to the robot.
+the robot coordinate system and the touchscreen is known
+(Fig.~\ref{fig|architecture}.13). And this robot location can also be used to
+identify gazes from the child to the robot.
 
 To make the children believe the robot is moving objects on the touchscreen, we
-synchronise a moving pointing gesture of the robot and a series of fake touches
-appied on the screen, moving the desired object. Once an object and a goal
-position have been selected, a planner generate a path for this image using the
+synchronise a moving pointing gesture of the robot
+(Fig.~\ref{fig|architecture}.4) and a series of fake touches
+(Fig.~\ref{fig|architecture}.8) appied on the screen, moving the desired object.
+Once an object and a goal position have been selected, a planner
+(Fig.~\ref{fig|architecture}.6) generate a path for this image using the
 A\mbox{*} algorithm on an occupancy map obtained with the items footprints, then
@@
 
-\paragraph{Wizard-of-Oz} To allow an experimenter to control the robot, a GUI to control the robot is
-provided and presents an identical representation of the state of the game on an
-other application which can be used on a tablet for example. The wizard can drag
-the objects in a similar fashion as what the child would do on the Sandtray, and
-on the release, the robot executes the dragging motion on the Sandtray, moving
-an object to a new location. The source code can easily modify to add new
-specific buttons to execute other actions, such as having the robot talk to the
-child.
+\paragraph{Wizard-of-Oz} To allow an experimenter to control the robot, a GUI to
+control the robot (Fig.~\ref{fig|architecture}.11) is provided and presents an
+identical representation of the state of the game on an other application which
+can be used on a tablet for example. The wizard can drag the objects in a
+similar fashion as what the child would do on the Sandtray, and on the release,
+the robot executes the dragging motion on the Sandtray, moving an object to a
+new location. The source code can easily modify to add new specific buttons to
+execute other actions, such as having the robot talk to the child.
 
@@
 
-A web interface can be used by the experimenter to manage the meta parameters of
-the interaction: log data on the subjects, start-stop, select a task, manage data recording...
+A web interface (Fig.~\ref{fig|architecture}.10) can be used by the experimenter
+to manage the meta parameters of the interaction: log data on the subjects,
+start-stop, select a task, manage data recording...
 

### 2017-10-06 paper.tex
Update on Overleaf.

@@
 
+We presented a set-up and data set of relatively unconstrained interaction between children and between a robot and a child.
+The set-up captures a rich set of multimodal streams which can be used to mine the social, verbal and non-verbal communication between two parties engaging in a relatively rich free play interaction.
+The data holds considerable promise for training social signal interpretation software, such as engagement interpretation or eye gaze reading.
+The dataset collected has sufficiently rich data and a wide range of multi-modal dimensions making it particularly suitable for Deep Learning of social signal processing algorithms.
+It also allow for very rich input to action selection mechanisms needed for autonomous robot behaviour.
+Future work will focus on using mining the data for patterns in free play, as per Parten's classification, and will attempt to extract social signals relevant to drive the interaction.
+Early results show that Deep Learning shows considerable promise for high-resolution tracking of eye gaze from the RGB video streams.
+
 \section*{Acknowledgments}

### 2017-10-06 paper.tex
Line wrappping

@@
 our dataset acquisition campaign (120 children, more than 45h of footage), the
-children's faces were automatically detected in 97\% \textcolor{red}{[TODO: check that!]} of the
-recorded frames.
+children's faces were automatically detected in 98\% of the recorded frames (see
+Section~\ref{sec|dataset}).
 
@@
 
-We presented a set-up and data set of relatively unconstrained interaction between children and between a robot and a child.
-The set-up captures a rich set of multimodal streams which can be used to mine the social, verbal and non-verbal communication between two parties engaging in a relatively rich free play interaction.
-The data holds considerable promise for training social signal interpretation software, such as engagement interpretation or eye gaze reading.
-The dataset collected has sufficiently rich data and a wide range of multi-modal dimensions making it particularly suitable for Deep Learning of social signal processing algorithms.
-It also allow for very rich input to action selection mechanisms needed for autonomous robot behaviour.
-Future work will focus on using mining the data for patterns in free play, as per Parten's classification, and will attempt to extract social signals relevant to drive the interaction.
-Early results show that Deep Learning shows considerable promise for high-resolution tracking of eye gaze from the RGB video streams.
+We presented a set-up and data set of relatively unconstrained interaction
+between children and between a robot and a child. The set-up captures a rich set
+of multimodal streams which can be used to mine the social, verbal and
+non-verbal communication between two parties engaging in a relatively rich free
+play interaction. The data holds considerable promise for training social signal
+interpretation software, such as engagement interpretation or eye gaze reading.
+The dataset collected has sufficiently rich data and a wide range of multi-modal
+dimensions making it particularly suitable for Deep Learning of social signal
+processing algorithms. It also allow for very rich input to action selection
+mechanisms needed for autonomous robot behaviour. Future work will focus on
+using mining the data for patterns in free play, as per Parten's classification,
+and will attempt to extract social signals relevant to drive the interaction.
+Early results show that Deep Learning shows considerable promise for
+high-resolution tracking of eye gaze from the RGB video streams.
 

### 2017-10-06 paper.tex
WIP on social annotations

@@
 
-\title{The Freeplay Sandbox: a Novel Methodology for the Evaluation of Social Robotics}
+\title{The Free-play Sandbox: a Novel Methodology for the Evaluation of Social Robotics}
 
@@
 
-The Freeplay Sandbox provides the opportunity to observe children interacting in a natural way in an open but framed setup.
+The Free-play Sandbox provides the opportunity to observe children interacting in a natural way in an open but framed setup.
 As the system can run on a single computer platform it can easily be deployed in the 'wild', in places where the children naturally interact such as classroom.
@@
 \label{ssec|CRI}
-This Freeplay Sandbox provides the opportunity to explore child-robot
+This Free-play Sandbox provides the opportunity to explore child-robot
 interactions in this open-real world environment as shown in Figure
@@
 with the manipulation/drawing \\ \midrule
-\bf Freeplay task (up to 40 min)\\
+\bf Free-play task (up to 40 min)\\
 \tabitem initial prompt: \emph{"Just to remind you, you can use the animals or draw. Whatever you
@@
 Annotating social interaction beyond surface behaviours is generally difficult:
-as research scientists, the behaviours we observe are typically the result of a
-superposition of complex and non-observable cognitive and emotional states. As
-such, these interesting socio-cognitive states can only be indirectly observed,
-and their labelling is error prone [TODO: citation needed].
+The observable, surface behaviours typically result of a superposition of the
+complex and non-observable underlying cognitive and emotional states. As
+such, these deeper socio-cognitive states can only be indirectly observed,
+and their labelling is typically error prone.
 
@@
 
-\subsection{Video Coding}
+In our scheme, we characterise \emph{Social engagement} by building upon
+Parten's stages of play. These 5 stages of play are normally
+used to characterise rather long sequences (at least several minutes) of social
+interactions.
 
-\todo{Mention the annotator tool}
+However, instead of characterising long social
+sequences (several minutes) with these 5 stages, we look at the micro-structure of the
+interaction:
+
+\paragraph{Social attitude}
+
+The constructs related to the social \emph{attitude} of the children derive from the
+\emph{Social Communication Coding System} (SCCS) proposed by
+Olswang~\etal\cite{olswang2006reliability}.  The SCCS consists in 6 mutually
+exclusive constructs characterising social communication (\emph{hostile};
+\emph{pro-social}; \emph{assertive}; \emph{passive}; \emph{adult seeking};
+\emph{irrelevant}) and were specifically created to characterise children
+communication in a classroom setting.
+
+We transpose these constructs from the communication domain to the general
+behavioural domain, keeping the \emph{pro-social}, \emph{hostile} (whose scope
+we broaden in \emph{adversarial}), \emph{assertive} (\ie dominant), and
+\emph{passive} constructs. In our scheme, the \emph{adult seeking} and
+\emph{irrelevant} constructs belong to Task Engagement axis.
+
+Finally, we have added the construct \emph{Frustrated} to describe children who
+are reluctant or refuse to engage in a specific phase of interaction
+because of a perceived lack of fairness or attention from their peer, or because
+they fail at achieving a particular task (like a drawing).
+
+\paragraph{Video coding}
+
+The coding is performed post-hoc with the help of a dedicated annotation tool
+(Fig.~\ref{annotator} which is part of the Free-play sandbox toolbox. This tool
+can replay and randomly seek in the three video streams, synchronised with the
+recorded state of the game (including the drawings as they are created). An
+interactive timeline displaying the annotations is also displayed.
+
+The annotation tool offers a remote interface for the annotator (made of large
+buttons, and visually similar to Figure~\ref{fig|coding-scheme}) that is
+typically displayed on a tablet and allow the simultaneous coding of the
+behaviours of the two children.  Usual video coding practices (double-coding of
+a portion of the dataset and calculation of an inter-judge agreement score)
+would have to be followed.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{annotator}
+    \caption{Screenshot of the dedicated tool developed for rapid annotation of
+    the social interactions.}
+    \label{annotator}
+\end{figure}
 
@@
 
-We have been using the Freeplay Sandbox task for an initial, large scale, data
+We have been using the free-play Sandbox task for an initial, large scale, data
 collection over a period of 3 months during Spring 2017.
@@
 were reliable (they were), and (3) establishing two experimental baselines for
-the Freeplay Sandbox task: the `human' baseline on one hand (child-child
+the Free-play Sandbox task: the `human' baseline on one hand (child-child
 condition), an `asocial' baseline on the other hand (child - \emph{non-social}
@@
 
-In the child-child condition (as seen in Figure~\ref{annotator}), 45 freeplay interactions (\ie 90 children) were
+In the child-child condition (as seen in Figure~\ref{annotator}), 45 free-play interactions (\ie 90 children) were
 recorded with a duration M=24.15 min (SD=11.25 min).
@@
 TODO: percentage faces seen?
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{annotator}
-    \caption{Screenshot of the dedicated tool developed for rapid annotation of
-    the social interactions.}
-    \label{annotator}
-\end{figure}
 

### 2017-10-06 paper.tex
Finished social annotations

@@
 
-Annotating social interaction beyond surface behaviours is generally difficult:
+Annotating social interaction beyond surface behaviours is generally difficult.
 The observable, surface behaviours typically result of a superposition of the
@@
 Our aim is to provide insights on the social dynamics, and we have synthesised a
-new coding scheme for social interactions that reuse and adapt three established
+new coding scheme for social interactions that reuse and adapt established
 social scales. Our coding scheme (Figure~\ref{fig|coding-scheme}) looks specifically
@@
 
+\paragraph{Task engagement}
+
+The first axis of our coding scheme aims at making a broad distinction between
+`on-task' behaviours (even tough the free-play sandbox does not explicitly
+require the children to perform a specific task, they are still engaged in an
+underlying task: to play with the game) and `off-task' behaviours. We call
+`on-task' behaviours \emph{goal oriented}: they encompass considered, planned
+actions (that might be social or not). \emph{Aimless} behaviours (with respect
+to the task) encompass opposite behaviours: being silly, chatting about
+unrelated matters, having a good laugh, etc. These \emph{Aimless} behaviours are
+in fact often highly social, and play an important role in establishing trust
+and cooperation between the peers. In that sense, they should not be discarded.
+
 \paragraph{Social engagement: Parten's stages of play at micro-level}
@@
 used to characterise rather long sequences (at least several minutes) of social
-interactions.
+interactions. Here, we apply them at the level of each of the micro-sequences of
+the interactions: one child is drawing and the other is observing is labelled as
+\emph{solitary play} for the former child, \emph{on-looker} behaviour for the
+later; the two children discuss what to do next: this sequence is annotated as
+a \emph{cooperative} behaviour; etc.
 
-However, instead of characterising long social
-sequences (several minutes) with these 5 stages, we look at the micro-structure of the
-interaction:
+By suggesting such a fine-grained coding of social engagement, we enable proper
+analyses of the internal dynamics of a long sequence of social interaction.
 

### 2017-10-06 paper.tex
Fixed caption Fig 4
Formatting

@@
 
-The Free-play Sandbox provides the opportunity to observe children interacting in a natural way in an open but framed setup.
+The free-play sandbox provides the opportunity to observe children interacting in a natural way in an open but framed setup.
 As the system can run on a single computer platform it can easily be deployed in the 'wild', in places where the children naturally interact such as classroom.
@@
 \label{ssec|CRI}
-This Free-play Sandbox provides the opportunity to explore child-robot
+This free-play sandbox provides the opportunity to explore child-robot
 interactions in this open-real world environment as shown in Figure
@@
 
-With the quantity of data logged and the high number of interaction achievable with the Freeplay Sandbox, it supports the type of requirement for recent Machine Learning approaches such as deep learning.
-The similar position of the children in all interactions makes the combination of data from different interaction easier than other less compact systems.
+With the quantity of data logged and the high number of interaction achievable
+with the free-play sandbox, it supports the type of requirement for recent
+Machine Learning approaches such as deep learning. The similar position of the
+children in all interactions makes the combination of data from different
+interaction easier than other less compact systems.
 
-From the information collected on the children, social behaviours can be extracted and used on a robot.
+From the information collected on the children, social behaviours can be
+extracted and used on a robot.
 
@@
     \includegraphics[width=\linewidth]{rviz-sandtray}
-    \caption{The free-play sandbox, viewed at runtime within ROS RViz. The
-    current background drawing (bottom left) is published on a regular image
-    topic. Simple computer vision is used to segment it into zones (visible in
-    the central panel). The poses and bounding boxes of the interactive items
-    are published as well, and turned into an occupancy map, used to plan the
-    robot's arm motion.}
+    \caption{The free-play sandbox, viewed at runtime within ROS RViz. Simple
+    computer vision is used to segment the background drawings into zones
+    (visible on the right panel). The poses and bounding boxes of the
+    interactive items are published as well, and turned into an occupancy map,
+    used to plan the robot's arm motion.}
+
     \label{fig|rviz}
@@
 We typically adhere to the acquisition procedure described in
-Table~\ref{tab|protocal} with all participants.  To ease later identification,
+Table~\ref{tab|protocol} with all participants.  To ease later identification,
 each child is also given a different and brightly coloured sports bib to wear.
@@
 
-As of today, no reliable Speech Recognition Engine exists for
+As of today, no reliable speech recognition engine exists for
 children~\cite{kennedy2017child}, but in the future, the audio should provide
@@
 The coding is performed post-hoc with the help of a dedicated annotation tool
-(Fig.~\ref{annotator} which is part of the Free-play sandbox toolbox. This tool
+(Fig.~\ref{annotator} which is part of the free-play sandbox toolbox. This tool
 can replay and randomly seek in the three video streams, synchronised with the
@@
 
-We have been using the free-play Sandbox task for an initial, large scale, data
+We have been using the free-play sandbox task for an initial, large scale, data
 collection over a period of 3 months during Spring 2017.
@@
 were reliable (they were), and (3) establishing two experimental baselines for
-the Free-play Sandbox task: the `human' baseline on one hand (child-child
+the free-play sandbox task: the `human' baseline on one hand (child-child
 condition), an `asocial' baseline on the other hand (child - \emph{non-social}

### 2017-10-06 paper.tex
Update dataset with paragraph duration

@@
 
-TODO: percentage faces seen?
+Over the dataset, the children faces are detected on 98\% of the images, which
+validates the location of the camera and the children to use the cameras to obtain
+facial social features.
 
@@
 \end{figure}
+
+\todo{Check paragrap}
+Figure \ref{fig|durations} presents an histogram of the durations of the
+interactions for the two conditions. In the child-child condition, we can
+observe two peaks, one around 10 to 15 minutes and one at 40 minutes, the
+maximum allowed by the protocol. This indicates that the task is engaging for
+the children and encourage them to play with each other for a consequent amount
+of time and provides opportunities to display the social behaviours we aimed to
+observe with it. In the child-robot condition, the peak at 40 minutes does not
+appear (3 children play alongside the robot until the end of the time
+allocated). The non-social robot appears to be less engaging than another child
+but still elicit plays of an average duration of 19.2 minutes which both is
+long enough to display social behaviours, and leaves us space for improvements
+if the robot policy is more social.
 

### 2017-10-06 paper.tex
Add back the description of the sandtray

@@
 
-The background image depicts a contextual environment, with different 
-zones (for example, water, grass, path, bushes...). Additionally, the children can draw on
-the background using a tool of the colour is costumisable. The players do
-not have any particular task to complete, they are simply invited to freely
-play. Importantly, they can play for as long as they wish (for
-practical reasons, we have limited the sessions to a maximum of 40 minutes in
-our own experiments, see Section~\ref{sec:dataset}).
+The free-play sandbox follows the sandtray
+paradigm~\cite{baxter2012touchscreen}: a large touchscreen (60cm $\times$ 33cm,
+with multitouch support) is used as an interactive surface (\emph{sandtray}).
+Two children play together by freely moving interactive items on the surface
+(Fig.~\ref{fig|sandbox}). A background image depicts a generic empty
+environment, with different symbolic colours (water, grass, beach, bushes...).
+By drawing on top of the background picture, the children can change the
+environment to their liking. The players do not have any particular task to
+complete, they are simply invited to freely play. Importantly, they can play for
+as long as they wish (for practical reasons, we have limited the sessions to a
+maximum of 40 minutes in our own experiments, see Section~\ref{sec:dataset}).
 

### 2017-10-07 paper.tex
Reworked discussion and conclusion

@@
 
-\section{Discussion}
+\section{Discussion \& Conclusion}
 \label{sec:discussion}
 
-\subsection{Analysis of the free play sandbox}
+\subsection{Analysis of the free-play sandbox}
 
-Our free play sandbox elicits a loosely structured play situation. The
-situation is indeed essentially aimless (free play). The goals and play episodes
-that we observe during the interactions are not pre-defined, and essentially
-unknown until they are decided by the players. In this sense, the free play
-sandbox permits and elicits what is fundamentally \emph{unstructured play}.
-\todo{Is it really aim-less? Children create their own goals, no?}
+The free play sandbox elicits a \emph{loosely structured} form of play: the
+actual play situations are not known and might change several times during the
+interaction; the game actions, even though based on a single interaction
+modality (the touchscreen), are varied and unlimited (especially when
+considering the drawings); the social interactions between participants
+are multi-modal (speech, body postures, gestures, facial expressions, etc.) and
+unconstrained. This loose structure creates a fecund environment for children to
+express a range of complex, dynamics, natural social behaviours that are not
+tied to an overly constructed social situation.
 
-However, a second look reveals some important structuring elements.
+The interaction is loosely structure. It is nonetheless structured:
+First, the physical bounds of the sandbox (an interactive table) limit the
+play area to a well defined and relatively small area. As a consequence,
+children are mostly static (they are sitting in front of the table) and their
+primary form of physical interaction is based on 2D manipulations on a screen.
 
-First, the physical bounds of the sandbox (an interactive table) limit the
-play zone to a well defined and relatively small area. As a consequence,
-children are mostly static (they are sitting in front of the table) and their
-primary form of interaction is 2D pick and place of items by drag and dropping
-them on the screen.
+Second, the game items themselves (visible in Figure~\ref{fig|sandbox})
+structure the game scenarios. They are iconic characters (animals or children)
+with strong semantics associated to them (like 'crocodiles like water and eat
+children'). The game background, with its recognizable zones, also elicit a
+particular type of games (like building a zoo or pretending we explore the
+savannah).
 
-Second, the game items themselves structure the game scenarios. Items are either
-iconic characters (animals or children) or plain blocks
-(Figure~\ref{fig|sandbox}). The former have strong semantics associated to them
-(like 'crocodiles like water and eat children'), while the later have little
-semantics associated. The game background, with its recognizable zones, also
-elicit a particular type of games (like building a zoo or pretending we explore
-the savannah).
+These elements of structure (along with other, less important, ones) make it
+possible for the free-play sandbox paradigm to retain some key properties that
+makes it a practical and effective scientific tool: because the game builds on
+simple and universal play mechanics (drawings, pretend play with characters),
+the paradigm is essentially cross-cultural; because the sandbox is physically
+bounded and relatively small, it can be easily transported and practically
+deployed in a range of environments (schools, exhibitions, etc.); because the
+whole apparatus is well defined and relatively easy to duplicate (it essentially
+consists in one single touchscreen computer), the free-play sandbox facilitates
+replication of findings in HRI while preserving ecological validity.
 
-Two more elements (the role and place of the experimenter, and the initial
-prompt given to the participant) have an important impact on the shape of the
-play situation. We discuss separately these two points in the description of the
-experimental procedure.
+%\paragraph{Forms of play elicited by the sandbox}
+%
+%\begin{itemize}
+%    \item dramatic play
+%    \item storytelling
+%    \item creative play
+%    \item pretend play
+%\end{itemize}
+%
+%\subsection{An exploration of the dynamics of social interactions}
+%
+%\begin{figure}
+%\centering
+%
+%\resizebox{0.8\linewidth}{!}{%
+%\begin{tikzpicture}[
+%                font=\sffamily,
+%                >=latex,
+%                every edge/.style={draw, very thick},
+%                node/.style={draw, rounded corners, align=center, inner sep=5pt, fill=black!20},
+%                label/.style={midway, align=center, font=\scriptsize\sffamily, fill=white}]
+%
+%    %%% NODES
+%    \node [node] (mentalrepresentations) {Mental Representations};
+%    \node [above left=of mentalrepresentations] (gameactions) {\it game actions};
+%    \node [node, above=of gameactions] (gamedynamics) {Game dynamics};
+%    \node [above right=of mentalrepresentations] (socialactions) {\it social actions};
+%    \node [node, above=of socialactions ] (socialdynamics) {Social Dynamics};
+%
+%    %%% CONNECTIONS
+%    \path (gamedynamics) edge [<->, dashed] node[label, above] {?}(socialdynamics);
+%    \path (gamedynamics) edge [<-, bend right] node[label] {mediate}(gameactions);
+%    \path (socialactions) edge [->, bend left] node[label] {reflect}(mentalrepresentations);
+%    \path (socialactions) edge [->, dashed] node[label,above] {?}(gamedynamics);
+%    \path (gameactions) edge [->, dashed] node[label,above] {?}(socialdynamics);
+%    \path (gameactions) edge [->, bend right] node[label] {reflect}(mentalrepresentations);
+%    \path (socialdynamics) edge [<-, bend left] node[label] {shape}(socialactions);
+%
+%\end{tikzpicture}
+%}
+%\label{fig|researchquestions}
+%\caption{Game dynamics, social dynamics and mental representations are linked
+%    with each other through social actions (implicit and explicit communication)
+%    and actions on the environment (mainly game actions)}
+%\end{figure}
+%
+%The mechanisms underlying the social interactions happening during free play
+%raise important questions for social robotics.
+%
+%
+%\paragraph{Game Dynamics}
+%
+%\begin{itemize}
+%    \item do we observe a sequence of sub-games?
+%    \item how are these sub-games initiated?
+%    \item how does A keep track of what B is playing at?
+%    \item how do we segment the action flow into meaningful play situations?
+%\end{itemize}
+%
+%\paragraph{Social Dynamics}
+%
+%\begin{itemize}
+%    \item does a ``social protocol'' establish?
+%    \item if so, what are these social rules?
+%    \item implicit vs explicit rule setting?
+%    \item what interaction modalities (or combination thereof) are relied upon
+%        to establish and maintain this ``social protocol''?
+%\end{itemize}
+%
+%\paragraph{Mental representations}
+%
+%\begin{itemize}
+%    \item Does A know what B is thinking of at time t?
+%    \item Do the participants perform explicit grounding of their respective mental models
+%        (ie, ask the other one what she is doing/thinking/planning to do?)
+%    \item How does the implicit grounding take place? attention tracking?
+%    \item Do we need at all to know what the other intends to do to actually
+%        collaborate? surface alignment vs deep grounding
+%    \item Procedural (how do you do it?) vs Semantic
+%        grounding (what are you doing? playing at?)
+%\end{itemize}
+%
+%\subsection{Stages of Play}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-Overall, the free play sandbox supports a \emph{loosely structured} form of play: the
-actual play situations are not known and might change several times during the
-interaction; the game actions, even though all based on a single modality (picking and
-placing game items), are unlimited; the social interactions between participants
-are multi-modal (speech, body postures, gestures, facial expressions, etc.) and
-unconstrainted. However, the broad domain of the play situations and the range of
-possible social interactions is bound by the physical bounds of the play zone
-and the theme of the game items.
-\todo{more than one actions, also drawing}
-
-\paragraph{Forms of play elicited by the sandbox}
-
-\begin{itemize}
-    \item dramatic play
-    \item storytelling
-    \item creative play
-    \item pretend play
-\end{itemize}
-
-\subsection{An exploration of the dynamics of social interactions}
-
-\begin{figure}
-\centering
-
-\resizebox{0.8\linewidth}{!}{%
-\begin{tikzpicture}[
-                font=\sffamily,
-                >=latex,
-                every edge/.style={draw, very thick},
-                node/.style={draw, rounded corners, align=center, inner sep=5pt, fill=black!20},
-                label/.style={midway, align=center, font=\scriptsize\sffamily, fill=white}]
-
-    %%% NODES
-    \node [node] (mentalrepresentations) {Mental Representations};
-    \node [above left=of mentalrepresentations] (gameactions) {\it game actions};
-    \node [node, above=of gameactions] (gamedynamics) {Game dynamics};
-    \node [above right=of mentalrepresentations] (socialactions) {\it social actions};
-    \node [node, above=of socialactions ] (socialdynamics) {Social Dynamics};
-
-    %%% CONNECTIONS
-    \path (gamedynamics) edge [<->, dashed] node[label, above] {?}(socialdynamics);
-    \path (gamedynamics) edge [<-, bend right] node[label] {mediate}(gameactions);
-    \path (socialactions) edge [->, bend left] node[label] {reflect}(mentalrepresentations);
-    \path (socialactions) edge [->, dashed] node[label,above] {?}(gamedynamics);
-    \path (gameactions) edge [->, dashed] node[label,above] {?}(socialdynamics);
-    \path (gameactions) edge [->, bend right] node[label] {reflect}(mentalrepresentations);
-    \path (socialdynamics) edge [<-, bend left] node[label] {shape}(socialactions);
-
-\end{tikzpicture}
-}
-\label{fig|researchquestions}
-\caption{Game dynamics, social dynamics and mental representations are linked
-    with each other through social actions (implicit and explicit communication)
-    and actions on the environment (mainly game actions)}
-\end{figure}
-
-The mechanisms underlying the social interactions happening during free play
-raise important questions for social robotics.
-
-\paragraph{Game Dynamics}
-
-\begin{itemize}
-    \item do we observe a sequence of sub-games?
-    \item how are these sub-games initiated?
-    \item how does A keep track of what B is playing at?
-    \item how do we segment the action flow into meaningful play situations?
-\end{itemize}
-
-\paragraph{Social Dynamics}
-
-\begin{itemize}
-    \item does a ``social protocol'' establish?
-    \item if so, what are these social rules?
-    \item implicit vs explicit rule setting?
-    \item what interaction modalities (or combination thereof) are relied upon
-        to establish and maintain this ``social protocol''?
-\end{itemize}
-
-\paragraph{Mental representations}
-
-\begin{itemize}
-    \item Does A know what B is thinking of at time t?
-    \item Do the participants perform explicit grounding of their respective mental models
-        (ie, ask the other one what she is doing/thinking/planning to do?)
-    \item How does the implicit grounding take place? attention tracking?
-    \item Do we need at all to know what the other intends to do to actually
-        collaborate? surface alignment vs deep grounding
-    \item Procedural (how do you do it?) vs Semantic
-        grounding (what are you doing? playing at?)
-\end{itemize}
-
-\subsection{Stages of Play}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Conclusion}
+\subsection{Towards the machine learning of social interactions?}
 \label{sec:conclusion}
@@
 of multimodal streams which can be used to mine the social, verbal and
-non-verbal communication between two parties engaging in a relatively rich free
-play interaction. The data holds considerable promise for training social signal
+non-verbal communication between two parties engaging in a rich free-play
+interaction. The data holds considerable promise for training social signal
 interpretation software, such as engagement interpretation or eye gaze reading.
@@
 mechanisms needed for autonomous robot behaviour. Future work will focus on
-using mining the data for patterns in free play, as per Parten's classification,
-and will attempt to extract social signals relevant to drive the interaction.
-Early results show that Deep Learning shows considerable promise for
-high-resolution tracking of eye gaze from the RGB video streams.
+mining the data for social patterns occurring in play situations, as per
+Parten's classification, and will attempt to extract social signals relevant to
+drive the interaction. Some early results show, for instance,  that deep
+learning shows considerable promise for high-resolution tracking of eye gaze
+from the RGB video streams.
 

### 2017-10-07 paper.tex
Finished missing bits -- submitted version

@@
 children's faces were automatically detected in 98\% of the recorded frames (see
-Section~\ref{sec|dataset}).
+Section~\ref{sec:dataset}).
 
@@
 
-\paragraph{Socio-cognitive framing of the task}
-
-\todo{Very much unfinished section.}
-focus on abstract socio-cognitive facets (robot
-perception and manipulation are simplified); well suited for qualitative and
-quantitative analysis using metrics like Słowinski’s Individual Motor Signature
-(for behavioural alignment), Anderson's~\cite{anderson2004social} coding of children’s free-play
-interactions, with-me-ness (for assessment of co-engagment).
-
-carefully framed: as far as possible, they necessitate only the cognitive
-capabilities that they evidence (e.g. if they evidence a purely socio-linguistic
-mechanism, they will not mandate complex action capabilities – they may benefit
-from it, though),
-
-This point is especially important as it allows an incremental path: a robot
-should be able to tackle one or several of these task independently of the
-others, and a research team may progressively extend their cognitive models to
-incorporate more and more of the cognitive capabilities required by the robot to
-address all of the tasks.
+%\paragraph{Socio-cognitive framing of the task}
+%
+%\todo{Very much unfinished section.}
+%focus on abstract socio-cognitive facets (robot
+%perception and manipulation are simplified); well suited for qualitative and
+%quantitative analysis using metrics like Słowinski’s Individual Motor Signature
+%(for behavioural alignment), Anderson's~\cite{anderson2004social} coding of children’s free-play
+%interactions, with-me-ness (for assessment of co-engagment).
+%
+%carefully framed: as far as possible, they necessitate only the cognitive
+%capabilities that they evidence (e.g. if they evidence a purely socio-linguistic
+%mechanism, they will not mandate complex action capabilities – they may benefit
+%from it, though),
+%
+%This point is especially important as it allows an incremental path: a robot
+%should be able to tackle one or several of these task independently of the
+%others, and a research team may progressively extend their cognitive models to
+%incorporate more and more of the cognitive capabilities required by the robot to
+%address all of the tasks.
 
@@
 \label{ssec|CRI}
+
 This free-play sandbox provides the opportunity to explore child-robot
-interactions in this open-real world environment as shown in Figure
+interactions in this open, real world environment as shown in Figure
 \ref{fig|freeplay}.
@@
 
-\section{Task Implementation}
+\section{Implementation}
 \label{sec:impl}
 
-The sandbox is implemented using mainly two frameworks: Qt's QML and
-\emph{QtQuick}\footnote{\url{http://doc.qt.io/qt-5/qmlapplications.html}}
-for the graphical interface of the game, and the \emph{Robot Operating System}
-(ROS) for the modular implementation of the data processing and behaviour
-generation pipelines. The graphical interface interacts with the decisional
-pipeline over a bidirectional QML-ROS bridge that we have developed for that
-purpose\footnote{Source code: \texttt{[hidden for blind review]}}.
+The software-side of the free-play sandbox is entirely
+open-source\footnote{Source code: \texttt{[hidden for blind review]}}. It is
+implemented using two main frameworks: Qt
+QML\footnote{\url{http://doc.qt.io/qt-5/qmlapplications.html}} for the graphical
+interface of the game, and the \emph{Robot Operating System} (ROS) for the
+modular implementation of the data processing and behaviour generation
+pipelines. The graphical interface interacts with the decisional pipeline over a
+bidirectional QML-ROS bridge that we have developed for that purpose.
 
@@
 layer between the background and the items, which adds another layer of
-unconstrained interaction to the game (eg. Figure~\ref{fig|sandbox}. The game
+unconstrained interaction to the game (Figure~\ref{fig|sandbox}). The game
 exposes the image of the background, the drawings, and the positions of the
-objects as a TF frame.
-
-\subsection{Sensors} Two RGB-D cameras are mounted at fixed positions on the
-Sandtray frame, each facing a child. This fixed position of the cameras allows
-the extraction of social features such as head pose, eye gaze, pointing and body
-posture, all described in a general coordinate system available to each software
-node. In addition to the image, the cameras posses a microphone to record the
-auditory aspect of the interaction. One node per camera
-(Fig.~\ref{fig|architecture}.2) outputs on dedicated topics the audio and video
-streams.
-
-\todo{briefly present ROS pipeline for sensors}
+objects as ROS TF frames.
+
+\subsection{Sensing} Two Intel RealSense SR300 RGB-D cameras are mounted at
+fixed positions on the sandtray frame, with custom designed 3D-printed brackets
+that ensure that the cameras are oriented towards the children's face. Because
+the cameras are rigidly mounted onto the sandtray's frame, their accurate
+geometric transformations with respect to the sandtray screen are known.
+Combined with hardware calibration, it allows for accurate localisation of the
+children and in particular, children's faces. In addition to the images, both
+cameras can perform stereo audio recording. One ROS node per camera
+(Fig.~\ref{fig|architecture}.2) publishes on dedicated topics the audio and video streams.
+
+A third `external' (and non-calibrated) camera is usually used as well to record
+the environment of the experiment with a wider angle (\emph{environment camera}
+in Figure~\ref{fig|freeplay}).
 
@@
 the robot executes the dragging motion on the Sandtray, moving an object to a
-new location. The source code can easily modify to add new specific buttons to
+new location. The source code can be easily modified to add new specific buttons to
 execute other actions, such as having the robot talk to the child.
 
-\todo{Should I add more in the WoZ GUI in the code side? It's fairly simple in the
-current implementation}
+%\todo{Should I add more in the WoZ GUI in the code side? It's fairly simple in the
+%current implementation}
 
@@
 
-A web interface (Fig.~\ref{fig|architecture}.10) can be used by the experimenter
-to manage the meta parameters of the interaction: log data on the subjects,
-start-stop, select a task, manage data recording...
+We have developed as well a dedicated, web-based, interface can be used by the experimenter
+to manage the whole experiment and data acquisition procedure
+(Fig.~\ref{fig|architecture}.10). This interface ensures that all the required
+software nodes are running, allow the experimenter to check the status and, if
+needed, to start/stop/restart any of them. It also help managing large data
+collection campaigns by providing a convenient web interface (usually used by
+the experimenter on a tablet) to record the demographics, resetting the game
+interface after each session, and automatically enforcing the acquisition
+protocol (see Table~\ref{tab|protocol}).
+
+This interface has been extensively used to acquire the dataset that we present at
+Section~\ref{sec:dataset}.
 
@@
 
-Online processing can be done on the images obtained from the camera. Using new
-deep learning libraries such as CMU's OpenPose~\cite{cao2017realtime} and
-DLib~(ADDREF), posture and facial landmarks can be obtained online. As the
-position of the camera, a potential robot and any object on the game is known,
-this landmarks can be mapped to high level behaviours such as poiting or looking
-at an object. Additional analysis can be done on the facial landmarks to other
-social states, such as main emotion felt by the child.
+Off-line post-processing can be done on the images obtained from the cameras. We
+rely on the CMU OpenPose library~\cite{cao2017realtime} to extract for both
+children the upper-body skeleton, 70 facial landmarks including the pupil
+position, as well as the hands' skeleton (when visible).
+
+Further processing is possible: As the position of the
+camera, a potential robot and any object on the game is known, this landmarks
+can be mapped to high level behaviours such as pointing or looking at an object.
+Additional analysis can be done on the facial landmarks to other social states,
+such as main emotion felt by the child.
 
@@
 
-\todo{check section title}
-
-Game features are also produced by the different nodes involved in the analaysis
+Game features are also produced by the different nodes involved in the analysis
 of the game. The Playground segmentation produce a map of the regions based on
@@
 regarding the methodology and the acquisition procedure can be found on the
-dataset website: \url{https://freeplay-sandbox.github.io/}. The dataset is open and accessible
-to any interested researcher.
-\todo{Probably add "an anonymised version of the dataser is open..." Reviewers might be concerned about ethics}
+dataset website\footnote{\url{https://freeplay-sandbox.github.io/}}. The dataset is open
+and accessible to any interested researcher, subject to adequate ethical clearance.
 
@@
 
-\todo{Check paragrap}
-Figure \ref{fig|durations} presents an histogram of the durations of the
-interactions for the two conditions. In the child-child condition, we can
-observe two peaks, one around 10 to 15 minutes and one at 40 minutes, the
-maximum allowed by the protocol. This indicates that the task is engaging for
-the children and encourage them to play with each other for a consequent amount
-of time and provides opportunities to display the social behaviours we aimed to
-observe with it. In the child-robot condition, the peak at 40 minutes does not
-appear (3 children play alongside the robot until the end of the time
-allocated). The non-social robot appears to be less engaging than another child
-but still elicit plays of an average duration of 19.2 minutes which both is
-long enough to display social behaviours, and leaves us space for improvements
-if the robot policy is more social.
+Figure~\ref{fig|durations} presents an histogram of the durations of the
+interactions for the two baselines. The distribution of the child-child
+interaction durations shows that (1) all children
+engage easily and for non-trivial amounts of time with the task; (2) the task
+leads to a wide range of level of commitment, which is desirable: it supports
+the claim that the free-play sandbox is an effective paradigm to observe a range
+of different social behaviours; (3) long interactions (>30 min) can result,
+which is especially desirable to study social dynamics.
+
+In contrast, and notwithstanding the smaller number of participants, the
+distribution of the child-robot interaction durations shows these interactions are
+generally shorter. This is expected as the robot was explicitly programmed not
+to interact with the children, resulting in a rather boring (and at time,
+awkward) situation where the child and the robot where playing side-by-side --
+in some case for rather long periods of time -- without interacting at all.
 
